Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-30 papers/thesis-kirstie/thesis.tex
edits

\documentclass[double,12pt]{beavtex}
\usepackage{graphicx}
\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command
\title{Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory}
\author{Kirstie L. Finster}
\degree{Master of Science}
\doctype{Thesis}
\department{Physics}
\depttype{Department}
\depthead{Head}
\major{Physics}
\advisor{David Roundy}
\submitdate{October ??, 2020}
\commencementyear{2020}

@@

\date{August 2020}

@@

The repulsive portion of the Leonard-Jones potential forms the potential for a 
Weeks-Chandler-Anderson (WCA) model fluid, shown in Figure \ref{fig:WCA_potential}. 
The WCA potential is described by the equation: 
\begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation} 
As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases 
from infinity as the distance between the centers of the atoms r increases, 
and becomes, and remains, essentially zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. 
As with the Lenard-Jones potential, the repulsive WCA potential increases 
rapidly and approches infinity as the two spheres begin to overlap. This 
is a more realistic repulsive potential than that of the Hard Sphere model 
which goes instantly to infinity when the two spheres make contact (at r=2R), 
and is zero otherwise. Because the WCA spheres have no hard boundary in 
their interatomic potential, they are called ``soft spheres'' and their 
gradually fading potential makes them slightly ``squishy''.

@@

The key idea behind classical Density Functional Theory (cDFT) is that 
free energy can be written as a functional (a function of a function) of 
a number density profile $n(\vec{r})$ which describes the way the number 
density, or atoms per volume, varies spatially on average. 
\begin{displaymath}{f[n(}\vec{r}{)]}\end{displaymath}
This, coupled with the fact that the free energy of a system tends toward 
a minimum at equilibrium means that the equilibrium free energy of the 
system ${f_{equil}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium 
number density profile $\rho{(}\vec{r}{)}$, can be found simply by varying 
$n(\vec{r})$ through all possible spatial profiles until the free energy 
is minimized.

@@

\section{Soft Fundamental Measure Theory (SFMT)}
In order to use cDFT, a functional for the excess Helmholtz free
energy for the particular fluid being modeled must be formed. But forming 
an expression based on Equation~\ref{eq:Fexcess-simplified} can be challenging. 
For the Hard Sphere fluid, there is another way to form an expression for 
the excess Helmholtz free energy in terms of a homogeneous number density 
and, that is to use Scaled Particle Theory (SPT).

Scaled Particle Theory seeks a general expression for the excess chemical 
potential of a homogeneous Hard Sphere fluid which can then be integrated 
to find the excess Helmholtz free energy.
%\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
It is an extension of the work done by Widom who showed that the excess 
chemical potential of a homogenesous fluid corredsponds to the reversible 
work of inserting a sphere of radius R into a fluid consisting of hard 
spheres of radius R. Widom's result was obtained by considering the difference 
in the excess Helmholtz free energy when one atom is added to an otherwise 
closed system. He showed that the excess chemcial potential is related to 
the ensemble average of the Boltzman factor for the potential energy 
difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
The resulting equation is known as Widom's Insertion Formula given by
%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
\color{red} \begin{equation}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3)\end{equation}
where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
hard sphere, R, and the cavity is either filled or not. \color{black}
%For cavities that are much larger than one of the hard spheres, 
Forming a more general result for different sizes of cavities in the fluid, 
Scaled Partical Theory arrives at an expression for the excess Helmholtz 
free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
\begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
where
\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}

An expression for the excess Helmholtz free energy as a functional of the 
number density profile $n(\vec r)$ can be found using Fundamental Measure 
Theory (FMT). FMT was developed for hard sphere fluids, and has been very 
sucessful at predicting the thermodynamic properties of inhomogeneous hard 
sphere fluids starting from an expression for the excess Helmholtz free 
energy obtained from Scaled Partical Theory. Soft Fundamental Measure 
Theory (SFMT), which will be addressed later, extends FMT to fluids with 
soft potentials like that of the WCA fluid.

@@

Incorporating into Equation~\ref{FexfromSPT} the two vector weights which 
were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
expression for $\Phi_3$ developed by Tarazona, the excess free energy becomes
\begin{displaymath}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}}\color{white}\right)}\end{displaymath}
\begin{equation}{\color{white}\left(\color{black}+\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}\right)\color{black}V}\end{equation} 
where the tensor weight is given by
\begin{equation}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
%and $\Phi_3$ is now
and
\begin{align}
\Phi_1 &= -n_{0}ln(1-n_{3}) \\
\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align}

@@

The equations developed so far have been for a hard sphere fluid. All of 
the weight function equations, as well as the equation for the deriviative 
of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists 
of a delta function characteristic of hard spheres. Soft Fundamental Measure 
Theory extends the Fundamental Measure Theory discussed so far to 
fluids with soft potenials by keeping the relationships between the weight 
functions, and that the weight functions satisfy the same, but replacing 
$w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
A new set of weight functions can then be derived for a system of soft spheres.

@@

\section{Overview}
%
%The goal of this project is to generate phase diagrams for a 
%Weeks-Chandler-Anderson model fluid using a computer program. 
%To do this we find the Helmholtz free energies that the fluid
%would have at a given temperature and density in a  
%liquid state, and in a crystalline solid state. 
%%Once the Helmholtz free energy for both the crystal and liquid states of 
%%the fluid are found for a given temperature and density, they can be 
%%compared to find the resulting state of the fluid. 
%Outside the region of 
%coexistence which exists between the liquid and crystal states, 
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature, 
%and construct phase diagrams.

@@

%\textcolor{red}{This cell is rather complicated for the purposes of this discussion, so 
%consider instead a simplified crystal cell with one atom in its center 
%as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
%minimized for a fixed volume (and temperature and number of particles).  
%A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
%be used for illustration. The number density n for a collection of 
%simplified cells is simply 1 atom per fixed bold-box volume.
%\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
%It is easier to think of 1 atom per ``box" (a reduced number density) 
%than 1 atom per some length$^3$ (the actual number density). Reduced number 
%density is given by $n^*$ and is equal to the number of atoms in a ``box" 
%of size $\sigma^3$.
%\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
%has a volume of $\sigma^3$, the reduced number density is just the number 
%of atoms in the fixed bold-box volume. 
%\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath} 
%}
In a large collection of cells, the atom at each lattice site of 
will not realistically reside exactly at the lattice point in the center 
of the cell. Statistically there will typically be a normal distribution 
of positions about the lattice point as pictured in the far right box in 
Figure \ref{fig:Ensemble_Gaus}. A plot of a 2-dimensional Gaussian 
distribution is shown in Figure \ref{fig:Gaus_plot}. 
A 3-dimensional Gaussian distribution centered at a lattice point can be used to represent 
a number density profile $n(\vec{r})$  which describes how the atoms are 
statistically distrubuted over a collection of cells. 
%In 3-dimensions, the Gaussian function takes on the form
%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int{e}^{-\frac{r^2}{2\sigma^2}}d\vec{r}}\end{equation} 
It is important to note that the number density does not change when introducing a number 
density profile. There are still the same number of atoms over some fixed volume. 
%There is still a total of one atom per ``bold box" volume.
%\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}
\begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
  \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
    \caption{A Gaussian Distribution Number Density Profile}
    \label{fig:Ensemble_Gaus}
  \end{figure}

@@

%The number density profile is varied by varying gw and fv, and the 
For a given combination of number density and temperature, a suite of number 
density profiles are generated.
% like those represented by Figure~\ref{fig:Ensemble_vary}.
The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number density profile,
and also for the homogeneous number denisty profile. 
Both the ideal and excess free energies are found, and added together to get the total Helmholtz 
free energy.
The Helmholtz free energy is minimized for a system with a fixed temperature, 
volume, and number of particles. Setting the number of particles to one, 
the volume can be found from the number density.
The process is then repeated to find the free energies at other combinations 
of number densities and temperatures.

@@

\section{Our Functional for the Excess Helmholtz free energy}
In order to apply classical Density Functional Theory, the Helmholtz energy
must be found. Computing the ideal portion of the Helmholtz free energy is
straightforward, but computing the excess portion of the Helmholtz free energy 
is difficult since it depends on the interatomic potential energy $V(r)$ between 
each sphere with all the other spheres. Soft Fundamental Measure Theory is a method
developed to form an expression for the excess Helmholtz free energy without
explicitly taking into account the interaction of each sphere with all the others, 
but it must be taylored to fit a particular fluid. 
%The excess Helmholtz free energy is particular to the fluid being modeled, 
%as it depends on the interatomic potential V(r).

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=9cm]{B2_vs_T_with_xi_fromB2.pdf}
    \caption{Shows the temperature dependence of $B_2$ generated from 
    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
    at each temperature.}
    \label{fig:B2vsT}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=9cm]{xi_fromB2.pdf}
    \caption{Shows the temperature dependence of $\Xi$ generated from 
    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
    at each temperature.}
    \label{fig:xi_fromB2vsT}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=9cm]{erf_Potential_xifromB2.pdf}
    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
    with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ 
    shown in black.}
    \label{fig:erf_Potential_xifromB2}
  \end{figure}

@@

\color{blue}
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath} 
\begin{equation}\label{compare_equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
\color{black}

>>>>

\documentclass[double,12pt]{beavtex}
\usepackage{graphicx}
\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command
\title{Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory}
\author{Kirstie L. Finster}
\degree{Master of Science}
\doctype{Thesis}
\department{Physics}
\depttype{Department}
\depthead{Head}
\major{Physics}
\advisor{David Roundy}
\submitdate{October ??, 2020}
\commencementyear{2021}

@@

\date{October 2020}

@@

The repulsive portion of the Leonard-Jones potential forms the potential for a 
Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}, 
shown in Figure \ref{fig:WCA_potential}. 
The WCA potential is described by the equation: 
\begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation} 
As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases 
from infinity as the distance between the centers of the atoms r increases, 
and becomes, and remains, essentially zero when 
r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. 
As with the Lenard-Jones potential, the repulsive WCA potential increases 
rapidly and approches infinity as the two spheres begin to overlap. This 
is a more realistic repulsive potential than that of the Hard Sphere model 
which goes instantly to infinity when the two spheres make contact (at r=2R), 
and is zero otherwise. Because the WCA spheres have no hard boundary in 
their interatomic potential, they are called ``soft spheres'' and their 
gradually fading potential makes them slightly ``squishy''.

@@

The key idea behind classical Density Functional Theory (cDFT) is that 
free energy can be written as a functional (a function of a function) of 
a number density profile $n(\vec{r})$ which describes the way the number 
density, or atoms per volume, varies spatially on average. 
\begin{displaymath}{f[n(}\vec{r}{)]}\end{displaymath}
This, coupled with the fact that the free energy of a system tends toward 
a minimum at equilibrium means that the equilibrium free energy of the 
system ${f_{equil}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium 
number density profile $\rho{(}\vec{r}{)}$, can be found simply by varying 
$n(\vec{r})$ through all possible spatial profiles until the free energy 
is minimized~\cite{MoritaDFT}.

@@

\section{Soft Fundamental Measure Theory (SFMT)}
In order to use cDFT, a functional for the excess Helmholtz free
energy for the particular fluid being modeled must be formed. But forming 
an expression based on Equation~\ref{eq:Fexcess-simplified} can be challenging. 
For the Hard Sphere fluid, there is another way to form an expression for 
the excess Helmholtz free energy in terms of a homogeneous number density 
and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.

Scaled Particle Theory seeks a general expression for the excess chemical 
potential of a homogeneous Hard Sphere fluid which can then be integrated 
to find the excess Helmholtz free energy.
%\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
It is an extension of the work done by Widom who showed that the excess 
chemical potential of a homogenesous fluid corredsponds to the reversible 
work of inserting a sphere of radius R into a fluid consisting of hard 
spheres of radius R. Widom's result was obtained by considering the difference 
in the excess Helmholtz free energy when one atom is added to an otherwise 
closed system. He showed that the excess chemcial potential is related to 
the ensemble average of the Boltzman factor for the potential energy 
difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
The resulting equation is known as Widom's Insertion Formula given by
%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
\color{red} \begin{equation}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3)\end{equation}
where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
hard sphere, R, and the cavity is either filled or not. \color{black}
%For cavities that are much larger than one of the hard spheres, 
Forming a more general result for different sizes of cavities in the fluid, 
Scaled Partical Theory arrives at an expression for the excess Helmholtz 
free energy of a homogeneous fluid~\cite{Hansen} in terms of the packing fraction $\eta$ 
\begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
where
\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}

An expression for the excess Helmholtz free energy as a functional of the 
number density profile $n(\vec r)$ can be found using Fundamental Measure 
Theory (FMT)~\cite{rosenfeld1989}. FMT was developed for hard sphere fluids, and has been very 
sucessful at predicting the thermodynamic properties of inhomogeneous hard 
sphere fluids starting from an expression for the excess Helmholtz free 
energy obtained from Scaled Partical Theory~\cite{santos2012phi3}. 
Soft Fundamental Measure Theory (SFMT), which will be addressed later, 
extends FMT to fluids with soft potentials like that of the WCA fluid.

@@

Incorporating into Equation~\ref{FexfromSPT} the two vector weights which 
were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
expression for $\Phi_3$ developed by Tarazona~\cite{tarazonaphi3, santos2012phi3}, 
the excess free energy becomes
\begin{displaymath}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}}\color{white}\right)}\end{displaymath}
\begin{equation}{\color{white}\left(\color{black}+\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}\right)\color{black}V}\end{equation} 
where the tensor weight is given by
\begin{equation}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
%and $\Phi_3$ is now
and
\begin{align}
\Phi_1 &= -n_{0}ln(1-n_{3}) \\
\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align}

@@

The equations developed so far have been for a hard sphere fluid. All of 
the weight function equations, as well as the equation for the deriviative 
of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists 
of a delta function characteristic of hard spheres. Soft Fundamental Measure 
Theory extends the Fundamental Measure Theory discussed so far to 
fluids with soft potenials by keeping the relationships between the weight 
functions, and which the weight functions satisfy the same, but replacing 
$w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
A new set of weight functions can then be derived for a system of soft 
spheres which is what we do in forming our functional for the 
Weeks-Chandler-Anderson fluid as detailed in the chapter on Methods.

@@

\section{The Overall Approach}
%
%The goal of this project is to generate phase diagrams for a 
%Weeks-Chandler-Anderson model fluid using a computer program. 
%To do this we find the Helmholtz free energies that the fluid
%would have at a given temperature and density in a  
%liquid state, and in a crystalline solid state. 
%%Once the Helmholtz free energy for both the crystal and liquid states of 
%%the fluid are found for a given temperature and density, they can be 
%%compared to find the resulting state of the fluid. 
%Outside the region of 
%coexistence which exists between the liquid and crystal states, 
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature, 
%and construct phase diagrams.

@@

%\textcolor{red}{This cell is rather complicated for the purposes of this discussion, so 
%consider instead a simplified crystal cell with one atom in its center 
%as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
%minimized for a fixed volume (and temperature and number of particles).  
%A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
%be used for illustration. The number density n for a collection of 
%simplified cells is simply 1 atom per fixed bold-box volume.
%\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
%It is easier to think of 1 atom per ``box" (a reduced number density) 
%than 1 atom per some length$^3$ (the actual number density). Reduced number 
%density is given by $n^*$ and is equal to the number of atoms in a ``box" 
%of size $\sigma^3$.
%\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
%has a volume of $\sigma^3$, the reduced number density is just the number 
%of atoms in the fixed bold-box volume. 
%\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath} 
%}
In a large collection of cells, the atom at each lattice site of 
will not realistically reside exactly at the lattice point in the center 
of the cell. Statistically there will typically be a normal distribution 
of positions about the lattice point as pictured in the far right box in 
Figure \ref{fig:Ensemble_Gaus}. A plot of a 2-dimensional Gaussian 
distribution is shown in Figure \ref{fig:Gaus_plot}. 
A 3-dimensional Gaussian distribution centered at a lattice point can be used to represent 
a number density profile $n(\vec{r})$  which describes how the atoms are 
statistically distrubuted over a collection of cells. 
%In 3-dimensions, the Gaussian function takes on the form
%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int{e}^{-\frac{r^2}{2\sigma^2}}d\vec{r}}\end{equation} 
It is important to note that the number density does not change when introducing a number 
density profile. There are still the same number of atoms over some fixed volume. 
%There is still a total of one atom per ``bold box" volume.
%\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}
\begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
where $n(\vec r)$ is itself an ensemble average over all microstates, and is independent of time:
\begin{equation}n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~\end{equation}

  \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
    \caption{A Gaussian Distribution Number Density Profile}
    \label{fig:Ensemble_Gaus}
  \end{figure}

@@

%Classical Density Functional Theory is used to find the equilbrium 
%Helmholtz free energy of the system at a given temperature and density
%by varying the number density profile $n(\vec{r})$ until the free energy 
%is minimized. The minimum free energy, and its corresponding number 
%density profile are then the equilibrium values for that temperature 
%and number density.

%The number density profile is varied by varying gw and fv, and the 
For a given combination of temperature and number density, a suite of number 
density profiles are generated similiar to those represented in the example 
set of profiles shown in Figure~\ref{fig:Ensemble_vary}. 
The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number 
density profile, and also for the homogeneous number denisty profile 
characteristic of a liquid. Both the ideal and excess free energies are 
found, and added together to get the total Helmholtz free energy.
According to classical Density Functional Theory, the minimum free energy, 
and its corresponding number density profile are then the equilibrium values 
for that temperature and number density.
%The Helmholtz free energy is minimized for a system with a fixed temperature, 
%volume, and number of particles. Setting the number of particles to one, 
%the volume can be found from the number density.
The process is repeated to find the free energies at other combinations 
of temperatures and number densities.

%Once the Helmholtz free energy for both the crystal and liquid states of 
%the fluid are found for a given temperature and density, they can be 
%compared to find the resulting state of the fluid. 
Outside the region of coexistence which exists between the liquid and 
crystal states, the fluid will be completely in a crystalline state if the 
crystal free energy is lower than the liquid free energy, and it will be 
completely in a liquid state if the liquid free energy is lower than 
the crystal free energy. Maxwell's Double Tangent Construction is used 
to determine the actual number densities of the crystal and liquid states
at the boundaries of the coexistance region for a given temperature, 
and construct phase diagrams as described in the section on The Theory of
Freezing and Maxwell's Double Tangent Construction.

@@

\section{Our Functional for the Excess Helmholtz free energy}
In order to apply classical Density Functional Theory, the Helmholtz energy
must be found. Computing the ideal portion of the Helmholtz free energy is
straightforward, but computing the excess portion of the Helmholtz free energy 
is difficult since it depends on the interatomic potential energy $V(r)$ between 
each sphere with all the other spheres. Soft Fundamental Measure Theory is a method
developed to form an expression for the excess Helmholtz free energy without
explicitly taking into account the interaction of each sphere with all the others, 
but it must be taylored to fit a particular fluid\cite{schmidt1999density}. 
%The excess Helmholtz free energy is particular to the fluid being modeled, 
%as it depends on the interatomic potential V(r).

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{B2_vs_T_with_xi_fromB2.pdf}
    \caption{Shows the temperature dependence of $B_2$ generated from 
    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
    at each temperature.}
    \label{fig:B2vsT}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{xi_fromB2.pdf}
    \caption{Shows the temperature dependence of $\Xi$ generated from 
    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
    at each temperature.}
    \label{fig:xi_fromB2vsT}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{erf_Potential_xifromB2.pdf}
    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
    with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ 
    shown in black.}
    \label{fig:erf_Potential_xifromB2}
  \end{figure}

@@

%\color{blue}
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath} 
\begin{equation}\label{compare_equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
%\color{black}
